Test_code

### 2016-08-22 code/HumanCdGF_microbiota_predict_by_day.Rmd
predict cfu by same and day 0 comm

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r, message=FALSE}
pack_used <- c('randomForest','miscTools','tidyr','ggplot2','dplyr', 'pROC', 'knitr', 'cowplot')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

@@

# read in files
meta_file   <- read.table('../data/process/human_CdGF_metadata.txt', sep = '\t',
                        header = T)
shared_file <- read.table('../data/process/human_CdGF.an.unique_list.0.03.subsample.shared',
                          sep='\t',header = T, row.names = 'Group')
tax_file <- read.table('../data/mothur/gf_cdiff.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy',
                       sep='\t',header = T, row.names = 'OTU')
# function to label otu with tax
source('tax_level.R')

# remove samples not used or with other cdiff strains
meta_file <- 
  meta_file %>% 
  filter(sample_type == 'stool', cdiff_strain == 431) %>% 
  select(Mouse_ID, group, day, cdiff_cfu, Euthanized) %>% 
  mutate(log_cfu = log10(cdiff_cfu + 1)) %>% 
  filter(!is.na(log_cfu))

@@

# create df with day 0 communities with column of mouse_id
day_0_otu_by_mouse <- 
  meta_file %>% 
  filter(day == 0 ) %>% 
  select(group, Mouse_ID) %>% 
  left_join(rel_abund_shared_01)

cfu_by_day <-
  meta_file %>% 
  select(group)

@@

RandomForest_by_day <- function(i, community){
if(community == day_0){
  rf_data <- 
  meta_file %>%
  filter(day == i) %>% 
  select(log_cfu, Mouse_ID) %>%
  inner_join(day_0_otu_by_mouse_01) %>% 
  select(-Mouse_ID)
} else if (community == by_day) {
  rf_data <- 
  meta_file %>%
  filter(day == i) %>% 
  select(log_cfu, Mouse_ID) %>%
  inner_join(day_0_otu_by_mouse_01) %>% 
  select(-Mouse_ID)

} else {
  print('Input for community needed (day_0 or by_day')
}

@@

predict_plot_day <-  
  ggplot(data = rf_prediction, aes(x=observed, y=predicted)) + geom_point() + 
        labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)') + 
        theme_bw() + ggtitle(paste("Random Forest Prediction of C Diff CFU \n on Day ",
                                   i, " from Day 0 Community", sep = '')) +
        annotate('text', x=0.15*(max(rf_prediction$observed)), 
                 y=0.95*(max(rf_prediction$predicted)), 
                 label=paste('R^2 = ', round(tail(rf_n_features$rsq, 1), 2),'\nMSE = ', 
                             round(tail(rf_n_features$mse, 1), 2), sep = ''))

@@

rf_plot_list[[i]] <- list(predict_plot_day,importance_plot_day)

})

rf_day_0_comm_plots <- lapply(1:10, day_0)
rf_by_day_com_plots <- lapply(1:10, by_day)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[1]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[1]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[2]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[2]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[3]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[3]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[4]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[4]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[5]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[5]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[6]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[6]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[7]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[7]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[8]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[8]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[9]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[9]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_plot_list[[10]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_plot_list[[10]][[2]], 0.5, 0, 0.5, 1)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
```

```{r}
pack_used <- c('randomForest','miscTools','tidyr','ggplot2','dplyr', 'pROC', 'knitr', 'cowplot')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

@@

# read in files
meta_file   <- read.table('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF_metadata.txt', sep = '\t',
                        header = T)
shared_file <- read.table('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/process/human_CdGF.an.unique_list.0.03.subsample.shared',
                          sep='\t',header = T, row.names = 'Group')
tax_file <- read.table('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/data/mothur/gf_cdiff.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy',
                       sep='\t',header = T, row.names = 'OTU')
# function to label otu with tax
source('~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/tax_level.R')

# identify c difficile otus
cdiff_otu <- rownames(tax_file[grep('Clostridium_XI', tax_file$Taxonomy),])
# remove samples not used or with other cdiff strains
meta_file <- 
  meta_file %>% 
  filter(sample_type == 'stool', cdiff_strain == 431) %>% 
  select(Mouse_ID, group, day, cdiff_cfu, Euthanized) %>% 
  mutate(log_cfu = log10(cdiff_cfu + 1)) %>% 
  filter(!is.na(log_cfu))

@@

# create df with day 0 communities with column of mouse_id
day_0_otu_by_mouse <- 
  meta_file %>% 
  filter(day == 0 ) %>% 
  select(group, Mouse_ID) %>% 
  left_join(rel_abund_shared_01) %>% 
  select(-group)
# create df with cfu merge to shared
cfu_by_day_comm <-
  meta_file %>% 
  filter(day %in% 1:10) %>% 
  select(group, day, log_cfu) %>% 
  left_join(rel_abund_shared_01) %>% 
  select(-group, -one_of(cdiff_otu))

@@

RandomForest_by_day <- function(i, community_day_0){
  if(!is.logical(community_day_0)){
    stop('Input for community  - Need community_day_0 = T/F')
  } else {
    rf_data <- if(community_day_0 == TRUE){
      meta_file %>%
        filter(day == 1) %>% 
        select(log_cfu, Mouse_ID) %>%
        inner_join(day_0_otu_by_mouse) %>% 
        select(-Mouse_ID)
    } else {
      cfu_by_day_comm %>% 
        filter(day == 1) %>% 
        select(-day)
    }

@@

    predict_plot_day <-  
      ggplot(data = rf_prediction, aes(x=observed, y=predicted)) + geom_point() + 
        labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)') + 
        theme_bw() + ggtitle(paste("Random Forest Prediction of C Diff CFU on Day ",
                                   i, sep = '')) +
        annotate('text', x=0.15*(max(rf_prediction$observed)), 
                 y=0.95*(max(rf_prediction$predicted)), 
                 label=paste('R^2 = ', round(tail(rf_n_features$rsq, 1), 2),'\nMSE = ', 
                             round(tail(rf_n_features$mse, 1), 2), sep = ''))

@@

    rf_plot_list[[i]] <- list(predict_plot_day,importance_plot_day)
  }
}
rf_day_0_comm_plots <- lapply(1:10, RandomForest_by_day, community_day_0 = T)
rf_by_day_com_plots <- lapply(1:10, RandomForest_by_day, community_day_0 = F)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[1]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[1]][[2]], 0.5, 0, 0.5, 1)

### Predict Day 1 CFU from Day 1 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[1]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[1]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[2]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[2]][[2]], 0.5, 0, 0.5, 1)

### Predict Day 2 CFU from Day 2 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[2]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[2]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[3]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[3]][[2]], 0.5, 0, 0.5, 1)

### Predict Day 3 CFU from Day 3 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[3]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[3]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[4]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[4]][[2]], 0.5, 0, 0.5, 1)

### Predict Day 4 CFU from Day 4 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[4]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[4]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[5]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[5]][[2]], 0.5, 0, 0.5, 1)

### Predict Day 5 CFU from Day 5 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[5]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[5]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[1]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[1]][[2]], 0.5, 0, 0.5, 1)

### Predict Day 6 CFU from Day 6 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[6]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[6]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[7]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[7]][[2]], 0.5, 0, 0.5, 1)

### Predict Day 7 CFU from Day 7 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[7]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[7]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[8]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[8]][[2]], 0.5, 0, 0.5, 1)

### Predict Day 8 CFU from Day 8 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[8]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[8]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[9]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[9]][[2]], 0.5, 0, 0.5, 1)

### Predict Day 9 CFU from Day 9 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[9]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[9]][[2]], 0.5, 0, 0.5, 1)

@@

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_day_0_comm_plots[[10]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_day_0_comm_plots[[10]][[2]], 0.5, 0, 0.5, 1)

@@

### Predict Day 10 CFU from Day 10 Community

```{r fig.height= 5, fig.width=10}
ggdraw()+
  draw_plot(rf_by_day_com_plots[[10]][[1]], 0, 0, 0.5, 1) + 
  draw_plot(rf_by_day_com_plots[[10]][[2]], 0.5, 0, 0.5, 1)

### Otus associated with CFU C difficile

```{r fig.width=10, fig.height=10}
predictive_otus <- paste('Otu00', c('199','026','050','015','028',
                                    '033','135','059','044'),    # lost with max rel abund > 10
                                    #'007', '014', '035', '005'), # gain with max rel abund > 10
                         sep = '')
cfu_by_day_comm %>% 
  select(log_cfu, one_of(predictive_otus)) %>% 
  filter(!is.na(log_cfu)) %>% 
  gather('otu','abundance',contains('Otu')) %>% 
  ggplot(aes(x = abundance, y = log_cfu)) + geom_point() + 
    facet_wrap(~ otu, scales = 'free_x') +
    labs( x = 'Relative Abundance', y = 'CFU (log10)')

```{r fig.width=10, fig.height=10}
cfu_by_day_comm %>% 
  select(log_cfu, one_of(predictive_otus)) %>% 
  filter(!is.na(log_cfu)) %>% 
  gather('otu','abundance',contains('Otu')) %>% 
  ggplot(aes(x = log10(abundance), y = log_cfu)) + geom_point() + 
    facet_wrap(~ otu, scales = 'free_x') +
    labs( x = 'Relative Abundance (log10)', y = 'CFU (log10)')
